---
title: "Election Violence Database Structure"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Election Violence Database Structure}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# devtools::install_github("bergant/datamodelr")
library(datamodelr)
```

This vignette describes the election violence database . 

## Overview of Database Structure

The election violence database contains information in four connected sections:

1.  **Event Report Coding**: The main section of the database contains all the project coding of reports of election violence in nineteenth and early twentieth century newspapers. Most users will be interested only (or primarily) in this part of the database. 

1.  **Source Material** The second section of the database contains information about the source material and search strategies used by the project to find that source material. 

1.  **Users** The third section of the database contains data about users.

1.  **Clustering** The fourth section of the database contains the data collected during the clustering process which grouped reports of the same event together.  

The diagram below shows the overall tabular structure of each section of the database. The diagram shows the primary and foreign key columns in each table, so the figure can serve as a reference for joining tables within and between sections. However, on first read through it is probably enough to note the broad relationship between the four sections of the database, and that there can be multiple tables in each section.


**Is this diagram too complex to go here? Could have a simple section level figure of the database here and put this diagram later (perhaps at the end).**  
*maybe put both complex diagram and simple section level figure here? That way users can focus on whichever they need (probably simple in forst instance, but they know where complex it when coming back to it)*
```{r, echo=FALSE, fig.height=7, fig.width=7}


dm <- datamodelr::dm_read_yaml("evp_erd.yml")
graph <- dm_create_graph(dm, view_type = "keys_only")
dm_render_graph(graph)
```

 

## Event Report Coding Tables

The main part of the database contains all the project coding of reports of election violence in nineteenth and early twentieth century newspapers. Most users will be interested only (or primarily) in this part of the database. The codings were generated by expert coders (users) reading newspaper articles (documents) and recording information about reports of election violence events reported in those newspaper articles, together with further classification of the event reports (tags and attributes). This coding is recorded in four related tables: 

1.  *user_docs* This table contains information about the allocation of users to newspaper articles. Each row in this table records the particular combination of user and document (allowing that users were responsible for coding multiple documents and one document could be coded by multiple different users).   
1.  **event_reports** This table stores information about reports of election violence. Each row in this table is a record of a report of an election violence incident (found in a particular newspaper article by a particular user). **This is probably the table in the database which users will want to understand first.** *agree* 
1.  *tags* This table stores further information about reported events like the location(s) where it took place, the actors involved in the event, and the kinds of action which constituted the event. Multiple tags of every kind could be associated with one reported event (e.g. an event could be associated with multiple locations, multiple actors and multiple actions).
1.  *attributes* The tags and attributes tables work together to store classification information about events. Basic classification information is stored in the tags table. The attributes table supplements the tags table with further more detailed information, where this information is associated with particular tags. This two layer structure enables flexible storage of classification information at different levels of granularity and in different ways. For example, if a tag identifies that the event report involes a violent disturbance, attributes might identify whether that this violent disturbance, was described as a riot in the press, had the riot act read and what the crowd size involved with the disturbance was.

The diagram below shows the fields in these four tables, and indicates how these four tables relate to each other. 

```{r, echo=FALSE, fig.height=7, fig.width=7}


graph <- dm_create_graph(dm, rankdir="RL", focus=list(tables = c("userdocumentallocation", "event_reports", "tag", "attribute")))
dm_render_graph(graph)
```

### user_docs (userdocumentallocation)
*user_doc_id*  
Unique identifying number for each event report allocated to a specific user.

*user_id*  
Unique user number - see users tables.

*document_id*  
Unique number for each newspaper article - see sources tables.

*allocation_date*  
Date at which the document was allocated to a user for coding. 

*allocation_type*  
Whether this allocation was for live coding purposes, or merely for training and testing of coding staff. Values of coding, training, testing.

*allocated_by*  
The nature and method of allocations - usually regular_random_assignment, in which documents were assigned randomly to two or more coders, with a 10% double-coding overlap. 

*coding_complete*  
Whether a coder finished categorising and article and submitted it.  1 or 0.

*article_type*  
What type of article the document? Coders could choose from the following:
 - Report of Trial or Court Proceedings. Usually summaries, with excerpts of evidence and speeches from lawyers and judges.
 - Report of Parliamentary Proceedings. Speeches made by MPs and Lords in the Houses of Parliament.
 - Letter(s) to Editor.Article takes the form of letters e.g. start with dear Sir, and end with the name of the writer.
 - Commentary, Editorial, Opinion-Piece. Generally marked by a more personal, opinionated tone, rather than a 'dispassionate' summary of events. Often make explicit moral judgements on events, with a clear political slant.
 - Printed Statement by Candidate or Party. Usually start with the preamble ‘To the Electors of…’, end with the name of the candidate, and take the form similar to that of a letter to the Editor.
 - Other.  Any type of article that does not correspond to those listed above. NOTE: The majority of articles coded fell into this category.

*geo_relevant*  
Does the article refer to event(s) in England and Wales? True or False.

*time_relevant*  
Does the article refer to event(s) which took place between 1832 and 1914? True or False

*electoral_nature*  
Does the article refer to events which occurred during an election? True or False.

*violence_nature*  
Does this article refer to violent events? True or False.

*electoralviolence_nature*  
Does the article contain information that may be of interest regarding the causes and consequences of electoral violence? True or False.  
NOTE: All questions (geo, time, electoral, violence, electoralviolence) must be TRUE to merit the creation of event reports.  Various combinations of TRUE and FALSE provide more precise knowledge of why an article chosen manually, or by our text classifier, was irrelevant (e.g. all TRUE, including electoralviolence, but FALSE for geo, usually indicates reports of Irish election violence).

*legibility*  
Whether the coder could read and understand the text of the article, either by clean-enough OCR or PDF image. True or False.

*comment_docinfo*  
Open-text qualitative comment by coder at doc info level, often indicating where there is uncertainty in the coding of other fields in the user_docs table, or the reason why an article is worth further qualitative investigation (see *recommend_qualitative*)

*status*  
Whether a coder finished categorising and article and submitted it, or is in progress. New, Saved, or Complete.

*recommend_qualitative*  
Qualitative judgement by the coder of whether, given the interests and aims of the project, the article contains information that may be of interest regarding the causes and consequences of electoral violence. True or False.

*difficulty_ranking*  
Qualitative judgement by the coder of how difficult the article was to read, categorise, and analyse. 

*ideal_coding_comments*  
Not in use. **Is this right?**

*score*  
Percentage matching score compared to double-codin of the same document by another user (where applicable - score of 0 where no double-coding exists).

*last_updated*  
That last date at which the article was amended and submitted on the codingp platform. 

*relevant*  
**Not sure, is it just 5 TRUE comments, or also not by-election, etc?**


### event_reports

This table stores information about reports of election violence. Each row in this table is a record of a report of an election violence incident (found in a particular newspaper article by a particular user).

*event_report_id*  
Unique identifying number for each event report.

*user_doc_id*  
See user_docs table.

*event_type*  
Field not in use.

*environment*  
Where the event took place - Outdoors, Indoors, Both, Unclear.

*event_start*  
The date at which an event began (or was estimated to begin).  Where this was not known, the default value was the date of newspaper publication (see documents table). Often same as event_end. Formatted as yyyy-mm-dd (e.g. 1868-11-24).

*event_end*  
The date at which an event ended (or was estimated to end).  Where this was not known, the default value was the date of newspaper publication (see documents table). Often same as event_start. Formatted as yyyy-mm-dd (e.g. 1868-11-24).

*comment_events*  
Open-text qualitative comment by coder, often indicating where there is uncertainty in the coding of other fields in the event reports table. 

*summary*  
Open-text qualitative summary of the election violence event. (e.g. 'riots after polling ended in Manchester. Candidate Lejeune attacked by the mob, escaped in carraige.  The Riot Act read and military brought in to restore order.') Usually between 1-3 sentences, but occasionally up to a paragraph.  

*meeting*  
Whether the event took place at a political meeting, including at the nomination/hustings, private political meetings indoors, public and outdoor meetings held by candidates or parties etc. - True or False or Unclear.

*election_point*  
At what point(s) in the election process did the event occur. Coders were able to select multiple options for events spanning multiple stages of the process.
 - Before Polling. This is during the canvass, or any other period explicitly stated to be before the commencement of the actual election. 
 - Nomination or Hustings. This will be explicitly referred to as nomination or hustings. It is likely to be mentioned alongside the existence of crowds, a platform, long speeches by candidates, and questions from the audience. 
  - During Polling. This is when voters were actually able to cast their votes for candidates. May only be alluded to (such as ‘during the election’), or such as when polling is suspended due to violence – this indicates that violence and the polling were simultaneous.
 - Declaration or Chairing. The point, often held in public and open-air settings, at which the winners and losers are announced. Occasionally, this is followed by the winning candidate being paraded through the town on a chair held aloft by supporters.
 - After Declaration or Chairing. Any point after the end of the formal election process.

*event_timeframe_quantifier*  
Related to start and end-dates.  Coders could select 'during' for a verified date range, or 'before' if they were unsure about the date of occurrence (e.g. an unspecified day before the default date of newspaper puublication).

*autodetected_cluster_id*  
See clustering tables. **this right?**

*is_exact*  
**Can't remember what this was** 

*latitude*  
The latitude coordinates of the event location.  For method of calculation, see *located_from*. 

*longitude*  
The latitude coordinates of the event location.  For method of calculation, see *located_from*. 

*election_id*  
Not in use.

*boundary_year*  
Constituency boundaries changed throughout the period covered by this database.  Constituency shapefiles and g_names were soured from Vision of Britain, and this field reflects the set of shapefiles related to the event date. Boundary year options are 1832, 1862, 1868, 1870, 1885. 

*byelection*  
A value of 0 indicates that an event relates to a General Election contest.  A value of 1 indicates that the event relates to a non-General Election contest.  This includes parliamentary by-elections, BUT ALSO, despite the field label, other contests such as local council elections, elections for churchwardens etc. 

*constituency_g_name*  
The name of the constituency related to the event.  Note that events which relate to a constituency do not always take place in that constituency, and will not be consistent with lag/long corrdinates.  This applies in cases where (e.g.) a county polling riot takes place within a borough constituency boundary because a polling station was located there, or events crossed boundaries at the edge of constituencies. These g_names and related stapefiles are sourced from Vision of Britain.

*county_g_name*  
The name of the county related to the event.  Should always be consistent with constituency_g_name, though some post-1885 constituency boundaries may overlap multiple counties. These g_names and related shapefiles are sourced from Vision of Britain.

*duration*  
Number of dates between event_start and event_end (usually 0)

*election_name*  
The General Election during which an event occurred.  This was automatically calculated using other fields, and should be ignored if a value is contained in this field, where byelection=1. 

*election_point_clean*  
Derived from election_point. Centres on polling - events are classed either as before polling, during polling, after polling, or unclear.

*end_weekday*  
The day of the week corresponding to event_end.

*geocluster*  
**Can't remember what this was**

*geometry_type*  
**Can't remember what this was**

*ignore_geounspecific*  
**Can't remember what this was**

*located_from*  
The method by which lat/long coordinates of an event were calculated.  This was done (e.g.) by matching to coder's town/village or constituency input [precise_tv and exact_constituency], or using the centrpoint of a county counstituency

*start_weekday*  
The day of the week corresponding to event_start.

*event_id*  
Not in use **is this right?**

### tag 
*Gary/Patrick we need a description of the table and the fields in the table here *
test

### attribute
*Gary/Patrick we need a description of the table and the fields in the table here *


## Sources Tables
The sources section of the database is designed to store information about the source documents which the reports of election violence are drawn from. Every event report is connected to a particular document (newspaper article). We located these documents by a search process. This search process involved:
1.Entering keywords into online newspaper archive search masks (overwhelmingly using the British Newspaper Archive)
1. Obtaining a set of results, that is the set of (links to) documents returned by the search.
1. Building the set of all newspaper articles returned by any search we conducted, with information about the article (these are the candidate documents - documents which may contain reference to election violence)
1. Selecting documents to code for instances of election violence.

The sources section of the database stores information about these documents, and connects these documents we found back through to the search process by which we found them. The information is stored in the following tables:

1. *archivesearches* This table stores information about the keyword searches we conducted.
1. *archivesearchresults* This table stores information about the results returned by the online archives in reponse to our keyword searches.
1. *candidatedocuments* This table indexes and table stores information about archive search results (so that a candidate document may appear in one or more rows in archivesearchresults).
1. *documents* This table stores information about the subset of candidate documents that were considered sufficiently likely to contain election violence incidents and were thus worth reading and coding. Every election violence report in the database is drawn from a document with an entry in this table. 

There is one additional table in the sources section of the database: *archivesearchsummaryonly*. This contains aggregate information (the number of results returned) about some searches. This information  does not connect to the event reports. Similar aggregate information, like the number of results returned, about searches which do connect to the event reports can be calculated from the archivesearchresults table.

```{r, echo=FALSE, fig.height=7, fig.width=7}


graph <- dm_create_graph(dm, focus=list(tables = c("documents", "candidatedocument", "archivesearchsummaryonly", "archivesearchresult", "archivesearch")), rankdir="RL")
dm_render_graph(graph)
```


**Gary/Patrick we need a table/field level description of each of these tables here**

## Users Tables
The users section of the database contains just one table which store information about database users. 
```{r, echo=FALSE}


graph <- dm_create_graph(dm, focus=list(tables = c("user")))
dm_render_graph(graph)
```


**Gary/Patrick we need a field level description of this table here**

## Clustering Tables

Most election violence events were reported many times in the nineteenth and early twentieth century newspapers, we have for example coded more than 150  newspaper articles reporting the 1857 election riot in Kidderminster. The clustering section of the database contains the information on the relationship between different event reports, and particularly which event reports are considered to be about the same underlying event. The clustering was done after all the event reports were collected. 

The idea of the clustering process is that every in scope^[Event reports are in scope if the coding is complete, the document is relevant, the event report relates to a general election, and if coding was undertaken in coding (as opposed to training or testing) modes. Out of scope event reports remain in the database but were not clustered.] event report should be a member of exactly one final cluster (combination cluster). The key information about any event report is its final (combination) cluster which is recorded in the combinedclusterentry table, which is linked back to event reports through the verifiedcluster and verifiedclusterentry table.

The database allows for this clustering process to be undertaken multiple times (different cluster attempts). Different cluster attempts are recorded in the clusterattempt_id column of the verified cluster table. The final cluster of an event report should therefore be unique *within* a particular cluster attempt.

The database contains stores information on multiple partial, and two major cluster attempts. **Gary should we add information here about the clusterattempts with ids?**. *I think so* The two main cluster attempts are 301:320, and 401:420 (covering 20 General Elections; 1832=301, 1835=302 etc.). All our our analysis of clustered data is based on attempts 401:420, which in addition to clustering on the platform, have undergone several manula post-clustering cleaning stages. 


These clusters were created a sequential grouping and comparison process:

1.  *Autoclustering:* Event reports with very similar features (e.g. similar locations in the same election) were put into clusters (autoclusters) by an automated algorithm
2.  *Verification* Coders compare all event reports in an autocluster, splitting as necessary to form a 'verified cluster'. 
3.  *Reallocation:* 'Orphan' event reports which were not part of a larger group at the autocluster stage were compared with all of these 'verified clusters', and reallocated to an existing or new verified cluster.
4.  *Combination* Verified clusters were compared to each other, and where verified clusters were determined to refere to the same underlying event, verified clusters were combined into larger final or 'combination clusters'. 

The clustering process was conducted within within election (i.e. event reports are only considered as possible matches to other event reports assigned to the same general election). Because of the sequential nature of the process (where information in tables was built up gradually by user entry so the information was often needed to processed by users in a partial state) some information is duplicated across tables. The authoritate route to the key information comes from: combinedclusterentry.combinedcluster_id, verifiedcluster.clusterattempt_id and verifiedclusterentry.event_report_id.

A full description of the clustering tables is found in the diagram below:

```{r, echo=FALSE, fig.height=7, fig.width=7}


graph <- dm_create_graph(dm, focus=list(tables = c("userreallocationallocation", "userautodetectedclusterallocation", "verifiedcluster", "verifiedclusterentry", "usercombinationallocation", "combinedclusterentry")))
dm_render_graph(graph)
```

**Gary/Patrick we need a table/field level description of each of these tables here**
