---
title: "Data on Searches of the Newspaper Archives"
author: "Gidon Cohen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data on Searches of the Newspaper Archives}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(durhamevp)
```

## Introduction

This vignette documents how the data from crawling the BNA and WNO archive is stored on the database, how this data can be retrived and how it can be analysed.

The database stores three information from the crawling process in three tables: 

  1.  The table portal\_archivesearch store information about the searches which have been run on the newspaper archives websites. 
  2.  The table portal\_archivesearchresults stores information about the results returned by these searches
  3.  The table portal\_candidatedocument stores the candidate documents


### Archive searches

The portal\_archivesearch table can be retrived with the `get_archivesearches()` command.

```{r getarchivesearches, eval=TRUE, warning=FALSE}

archivesearches<-get_archivesearches()

knitr::kable(tail(archivesearches, 3))

```


### Archive search results

The portal\_archivesearchresults table can be retrived with the `get_archivesearchresults()` command. To retrieve the results relating to a particular search. For example to retrieve the results returned by the most recent document search (archive_search_id = `r tail(archivesearches$id,1)`):


```{r getarchivesearchresults, eval=TRUE, warning=FALSE}
lastid<-tail(archivesearches$id,1)
archivesearchresults<-get_archivesearchresults(archive_search_id=lastid)

names(archivesearchresults)

dim(archivesearchresults)

knitr::kable(head(archivesearchresults[,c("id", "archive_search_id", "description", "publication_title")], 3))

```


